---
title: "Analysis"
author: "Team Student"
header-includes:
  - \usepackage{amsmath}
  - \usepackage{algorithm}
  - \usepackage{algorithmicx}
  - \usepackage[noend]{algpseudocode}
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This package is part of a project to determine the best way to evaluate data with rounded values (binned data).  Our task was to use an EM algorithm with penalty for mixtures of t-distributions.

This vignette has the goal to explain all important functions contained in this package and their usage as well as the visualization with examples.


```{r, include=F}
library(t.mix)
```




```{r}
data("ZD", package = "EUCASTData")
```


First we establish three example data sets, which will be used throughout this vignette. These datasets are subsets of our original data.

```{r}
#extract subsets
ZDs1 <- subset(ZD, Antimicrobial == "Ampicillin" & Bacterium == "Escherichia coli")
ZDs2 <- subset(ZD, Antimicrobial == "Piperacillin" & Bacterium == "Escherichia coli")
ZDs3 <- subset(ZD, Antimicrobial == "Mecillinam" & Bacterium == "Escherichia coli")
```


## Main functions for analysis
To get an idea of how our functions work, we provided an overview of our algorithm, which should help to understand the functions and their purpose.

\begin{algorithm}[H]
\caption{EM Algorithm for mixtures of t's with binned data (grouped)}\label{euclid}
\begin{algorithmic}[1]
\Procedure{}{}
\State Calculate starting values $\pi_k^{(0)}, \mu_k^{(0)}, \sigma_k^{(0)}, \nu_k^{(0)}$, \; $k=1, ..., K$
		\State Set $\alpha_0$, $\beta_0$
		
		
		\Repeat
		\State $\tau_{k, j1}^{(i)} = \frac{\pi_k^{(i)} f_{\tilde Y}(j|\mu_k^{(i)}, \sigma_k^{(i)},\nu_k^{(i)})}{\sum \limits_{c=1}^K \pi_c^{(i)} f_{\tilde Y}(j|\mu_c^{(i)}, \sigma_c^{(i)}, \nu_c^{(i)})}$ for $k=1, ..., K$, $j = 1, ..., J$
		\State $\pi_k^{(i+1)} \gets\frac{1}{\tilde N} \sum\limits_{j=1}^{J}  n_j \tau_{k, j1}^{(i)} \qquad \text{for } k=1, ..., K$
		
		\State $(\mu_k^{(i+1)}, \sigma_k^{(i+1)}, \nu_k^{(i+1)}) \gets \arg \max Q_2(\mu,\; \sigma, \; \nu;\;\Theta^{(i)}) + \newline Q_3(\sigma; \; \alpha_0, \beta_0) $ for $k=1, ..., K$
		\Until{convergence criterion is fulfilled}
		\State \Return $\Theta^{(i+1)}$
		
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

The two main functions to analyse the data are \textbf{total.function} and \textbf{model.select.t.mixture}. Most of the other functions included in this package do the preliminary work or are used in either one or the other or both of the main functions.

We will now proceed with evaluating the first example dataset.

```{r}
####### First data set #########

#extract data necessary for analysis

j = 6:50 #these values have been observed
n1 = as.integer(ZDs1[4:48]) #observed frequencies

#one component
(result1 = total.function(n1, j, K=1, atoms=6, memb.exp=2, draw=TRUE))

#two components
(result1 = total.function(n1, j, K=2, atoms=6, memb.exp=2, draw=TRUE))

#three components
(result1 = total.function(n1, j, K=3, atoms=6, memb.exp=2, draw=TRUE))
```

The last step is the model selection process:

```{r}
#model selection (up to Kmax=5)
(selected1 <- model.select.t.mixture(n1, j, Kmax=5, memb.exp=2, atoms=6))
```

This process is repeated for the second example dataset :

```{r}
####### Second data set ##########

n2 = as.integer(ZDs2[4:48])

#one component
(result2 = total.function(n2, j, K=1, atoms=6, memb.exp=2, draw=TRUE))

#two components
(result2 = total.function(n2, j, K=2, atoms=6, memb.exp=2, draw=TRUE))

#three components
(result2 = total.function(n2, j, K=3, atoms=6, memb.exp=2, draw=TRUE))
```


```{r}
#model selection (up to Kmax=5)
(selected2 <- model.select.t.mixture(n2, j, Kmax=5, memb.exp=2, atoms=6))
```

On to the third example dataset:

```{r}
###### Third data set #############


n3 = as.integer(ZDs3[4:48])

#one component
(result3 = total.function(n3, j, K=1, atoms=6, memb.exp=2, draw=TRUE))

#two components
(result3 = total.function(n3, j, K=2, atoms=6, memb.exp=2, draw=TRUE))

#three components
(result3 = total.function(n3, j, K=3, atoms=6, memb.exp=2, draw=TRUE))

#four components
(result3 = total.function(n3, j, K=4, atoms=6, memb.exp=2, draw=TRUE))
```

```{r}
#model selection (up to Kmax=5)
(selected3 <- model.select.t.mixture(n3, j, Kmax=5, memb.exp=2, atoms=6, optim.method="Nelder-Mead"))

```
